Outbreak response models of vaccine-preventable diseases in humans (1970-2019) - A systematic review

### 2023-06-25 analyses/scripts/orv_sys_rev_data_analysis.Rmd
Added more break downs of the collaboration type analyses

@@

>>>>

### Number of studies per year and broken down by collaboration types

```{r echo=FALSE}
#' Count the number of each author affiliation type per year (including FMD)
collap_types_by_year <- compact_data_with_citation_keys_no_fmd %>% 
  mutate(year_aggreg = as.numeric(ifelse(publication_year < 2006, 
                                         2005, 
                                         publication_year
                                         )
                                  ), 
         collab_type = as_factor(if_else(author_affiliation_type == 'academic_institutions', 
                                         'Purely academic', 
                                         'Mixed')
                                 )
         ) %>% 
  count(year_aggreg, collab_type, name = 'num_of_studies') %>% 
  pivot_wider(names_from = collab_type, values_from = num_of_studies) %>% 
  clean_names() %>% 
  adorn_totals(... = c(mixed, purely_academic), where = c("row", "col"))

knitr::kable(collap_types_by_year,
             caption = "Breakdown of studies per year by collaboration types")
```

```{r echo=FALSE}
#' Count by collaboration types and make a bar plot
collab_types_count <- compact_data_with_citation_keys_no_fmd %>%
  count(author_affiliation_type, sort = TRUE, name = "num_of_studies") %>% 
  mutate(author_affiliation_type = as_factor(author_affiliation_type)) %>% 
  arrange(desc(num_of_studies)) %>% 
  adorn_totals() %>% 
  adorn_percentages(denominator = 'col') %>% 
  adorn_pct_formatting() %>% 
  adorn_ns(position = "front")

knitr::kable(collab_types_count)
```

### Number of publications per year by collaboration type

```{r} 
 #' Count the number of each author affiliation type per year
 collabs_per_year <- compact_data_with_citation_keys_no_fmd %>%
   count(year_aggreg,
     collab_type,
     name = "total_publications"
   )
 x_axis_labels <- as.character(seq(2005, 2020, 2))

 collabs_per_year_plot <- ggplot(
   data = collabs_per_year,
   aes(
     x = year_aggreg,
     y = total_publications
   ),
   color = "black"
 ) +
   geom_bar(aes(fill = collab_type),
     stat = "identity",
     position = "dodge"
   ) +
   scale_fill_manual(values = c("tan3", "turquoise2")) +
   scale_x_continuous(
     breaks = seq(2005, 2020, 2),
     labels = x_axis_labels,
     expand = c(0, 0)
   ) +
   scale_y_continuous(
     breaks = seq(0, 20, 2),
     labels = seq(0, 20, 2)
   ) +
   labs(
     x = "Year",
     y = "Total publications",
     fill = "Collaboration type"
   ) +
   theme_minimal(base_size = 12) +
   theme(legend.position = "bottom")

 ggsave(
   filename = "total_collabs_per_year_plot.tiff",
   plot = collabs_per_year_plot,
   path = gh_figs_dir
 )

 ggsave(
   filename = "total_collabs_per_year_plot.tiff",
   plot = collabs_per_year_plot,
   path = dropbox_figs_dir
 )
```

Even though non-academic collaborations increased over time, the share of 
total publications per year over time decreased.

### Proportion of the total publications per year by collaboration type

```{r echo=FALSE, fig.width=8} 
#' Count the number of each author affiliation type per year (including FMD) 
collap_types_by_year <- compact_data_with_citation_keys_no_fmd %>%
  count(year_aggreg,
    collab_type,
    name = "total_publications"
  )

# The plot
# Modify the x-axis labels to account for the lumping of the years from 2005 and below
x_axis_labels <- as.character(seq(2005, 2020, 2))

x_axis_labels[1] <- "1970-2005"

collab_proportions_plot <- ggplot(data = collap_types_by_year) +
  geom_bar(
    aes(
      x = year_aggreg,
      y = total_publications,
      fill = collab_type
    ),
    color = "black",
    stat = "identity",
    position = position_fill(reverse = TRUE) # ,
    # show.legend = FALSE
  ) +
  scale_fill_manual(values = c("tan3", "turquoise2")) +
  scale_x_continuous(
    breaks = seq(2005, 2020, 2),
    labels = x_axis_labels,
    expand = c(0, 0)
  ) +
  labs(
    x = "Year",
    y = "Proportion of total publications",
    fill = "Collaboration type"
  ) +
  theme(legend.position = "none") +
  theme_minimal(base_size = 12)

plot(collab_proportions_plot)

ggsave(
  filename = "collab_proportions.tiff",
  plot = collab_proportions_plot,
  path = gh_figs_dir
)

ggsave(
  filename = "collab_proportions.tiff",
  plot = collab_proportions_plot,
  path = dropbox_figs_dir
)
```

### Mixed collaborations with and without academics

```{r echo=FALSE}
academic_collabs_versus_others <- compact_data_with_citation_keys_no_fmd %>%
  filter(author_affiliation_type != "academic_institutions") %>%
  mutate(
    year_aggreg = as.numeric(ifelse(publication_year < 2006, 2005, publication_year)),
    collab_type = as_factor(if_else(str_detect(
      author_affiliation_type,
      "academic_institutions"
    ),
    "including_academics",
    "excluding_academics"
    ))
  ) %>%
  count(year_aggreg, collab_type, name = "num_of_studies")

# The plot
# Modify the x-axis labels to account for the lumping of the years from 2005 and below
x_axis_labels <- as.character(seq(2005, 2020, 2))

x_axis_labels[1] <- "1970-2005"

# plot the result
academic_collabs_trend <- ggplot(data = academic_collabs_versus_others) +
  geom_bar(
    aes(
      x = year_aggreg,
      y = num_of_studies,
      fill = collab_type
    ),
    color = "black",
    stat = "identity",
    position = position_fill(reverse = TRUE)
  ) +
  scale_fill_manual(values = c("tan1", "turquoise1")) +
  scale_x_continuous(
    breaks = seq(2005, 2020, 2),
    labels = x_axis_labels
  ) +
  coord_flip() +
  labs(
    title = "Proportion of mixed collaborations per year",
    x = "Year",
    y = "Proportion of publications",
    fill = "Mixed collaboration type"
  ) +
  theme_minimal(base_size = 12)

plot(academic_collabs_trend)
```

@@

## Diseases and interventions

```{r echo=FALSE}
#count the number of unique interventions modelled
interventions_unique_entries <- long_data_no_fmd %>% 
  group_by(title) %>% 
  distinct(intervention_modelled, .keep_all = TRUE) %>% 
  ungroup()

#interventions are disease specific
interventions_per_disease <- interventions_unique_entries %>% 
  mutate(intervention_modelled = str_trim(intervention_modelled)) %>% 
  filter(intervention_modelled != 'e.g.') %>% #remove false rows
  filter(intervention_modelled != 'various') %>% #remove false rows
  count(disease, intervention_modelled, name = 'num_of_studies') %>% 
  arrange(disease, desc(num_of_studies)) %>% 
  ungroup()

knitr::kable(interventions_per_disease)
```
### Diseases and interventions (top 2 per disease)

```{r echo=FALSE}
#count the number of unique interventions modelled
interventions_per_disease_top2 <- long_data_no_fmd %>% 
  group_by(title) %>% 
  distinct(intervention_modelled, .keep_all = TRUE) %>% 
  ungroup() %>% 
  count(disease, intervention_modelled, name = 'num_of_studies') %>% 
  arrange(disease, desc(num_of_studies)) %>% 
  group_by(disease) %>% 
  slice_head(n = 2) %>% #keep only top 2 per group 
  ungroup()

knitr::kable(interventions_per_disease_top2,
             caption = "Top 2 most studied intervention per disease")
```
